### 2015-07-07 narrative/data-reduction.tex

\item Precision based re-organization. This is where the most significant bytes
	of the data are all group together from each object. This data
	generally will have a higher utility then the data with the least
	significant bytes.  This process has the following three steps which
	need to be efficiently implemented when data gets placed to the storage
	system and from the storage system.  First, the data needs to be
	re-arranged which involves memory operations and needs to be done
	in-situ. This operation involves no communication.  Data needs to be
	copied, so the memory requirements are increased. At most this requires
	two copies an individual dataset in memory, and we will investigate
	techniques to allocate and deallocate this memory if the user will
	specify that the data will be overwritten after it is written to the
	storage system. This is often the case for many of the data quantities
	written form the simulation, but there are many cases where, for
	example, we want to write all of the particles from a Particle In Cell
	(PIC) simulation.  Since the particles will be used later in the
	calculation, we need to duplicate the storage. Our observation with
	working with the XGC1, GTC, Warp, and PiconGPU simulations (all
	leadership class simulations) is that we can temporary increase the
	storage of the particles, and then release them since the temporary
	arrays used in the calculations are often freed when a PIC iteration is
	finished. 
%
\item Frequency-based re-organization.  Another common technique to classify
	the importance of data is to first re-organize the data in frequency
	space and then specify that as the frequency increases the importance
	of data decreases.   Common technique which is used for streaming data,
	and data reduction techniques such as JPEG-2000~\cite{jpeg2000} is to
	code streams which have regions of interest that offer several
	mechanisms to support spatial random access or region of interest
	access at varying degrees of granularity. It is possible to store
	different parts of the same data using different quality.  This allows
	us to place the lowest frequency pieces in the fastest storage, and the
	highest frequency either on the slowest storage tiers, or if the data
	sizes are prohibitively  costly, not even write out the highest
	frequency. This binning of data in frequency space can take as an input
	(storage bandwidth knowledge, storage size limitations, and user
	intentions of how long the majority of the data will be frequently
	accessed.  In order to fully take advantage of frequency based
	re-organization we have devised a scheme where the data pre-conditioner
	first sorts the data in bins of 1024 words, and then we maintain the
	sort order, and then we used wavelets and spline fits to then
	re-organize the values.  This re-organization allows us to understand
	that the resultant data is smoother now, and can be reduced in each one
	of the bins.  \end{itemize}

@@

We will also research the time it takes to re-organize the data and use the
multi-layer approach to staging to understand where the data gets reduced and
where it gets re-organized. Open questions are: 1) do we move the data over to
other nodes to then reduce and re-organize, 2) do we first reduce the data and
then later re-organize this data on different nodes?  3) do we use different
cores on a node to do data-refactoring.  Similarly we will examine the cost of
reading the data. It is clear that if we only read in the high-importance data
it should improve the overall time to move the data from the SSIO layer to the
application, but what happens when users want all of the data? There is an
obvious penalty of having to take data in a hierarchical layout and then place
this back to a single-layer view.  Additional research questions which must be
addressed: Will some types of data (PIC data, finite element data, ...) be be
conducive  for data-refactoring than others? How much semantic-domain knowledge
will we need to get a dramatic improvement of the storage system over not doing
anything special?

@@

\paragraph{High Impact Challenges:} Maximal data reduction on the SSIO layer
would result from storing only the code and the input parameters, and simply
recalculating all of the results. This extreme is not feasible in practice
owingto the size of the computation involved, but sheds light on the path to
achieving a step change in data reduction. We shall investigate the possibility
of writing a minimal set of information, and placing an ``auditor'' code in a
docker container so that data can be re-generated with a {\em local}
approximation of the full simulation. This is an entirely new approach to
tackle the data overload by using preprocessors in which information is not
simply stored in the form of raw data, but rather in the form of appropriate
parameterized algorithms. A priori information on the nature of the data is
used to choose the basic algorithm while the actual values of the data are used
to tune the parameters used in the reconstruction procedures. We shall refer to
this paradigm as an Auditor.  We will examine the use of several auditors for
the simulations which we will work with. One approach is to use a fluid-code to
audit a kinetic code, which can be a good approximation over short periods of
time. Once there is a large deviation (specified by the user) from the auditor
to the original calculation, the data will be saved. This is a local operation,
since the deviations will grow differently in space and in time.

>>>>

\item Precision based re-organization. This is where the most significant bytes
	of the data are all grouped together from each object. This data
	generally will have a higher utility then the data with the least
	significant bytes.  This process has the following three steps which
	need to be efficiently implemented when data gets placed to the storage
	system and from the storage system.  First, the data needs to be
	re-arranged which involves memory operations and needs to be done
	in-situ. This operation involves no communication.  Data needs to be
	copied, so the memory requirements are increased. At most this requires
	two copies an individual dataset in memory, and we will investigate
	techniques to allocate and deallocate this memory if the user will
	specify that the data will be overwritten after it is written to the
	storage system. This is often the case for many of the data quantities
	written from the simulation, but there are many cases where, for
	example, we want to write all of the particles from a Particle In Cell
	(PIC) simulation.  Since the particles will be used later in the
	calculation, we need to duplicate the storage. Our observation with
	working with the XGC1, GTC, Warp, and PiconGPU simulations (all
	leadership class simulations) is that we can temporary increase the
	storage of the particles, and then release them since the temporary
	arrays used in the calculations are often freed when a PIC iteration is
	finished. 
%
\item Frequency-based re-organization.  Another common technique to classify
	the importance of data is to first re-organize the data in frequency
	space and then specify that as the frequency increases the importance
	of data decreases.   Common technique which is used for streaming data,
	and data reduction techniques such as JPEG-2000~\cite{jpeg2000} is to
	code streams which have regions of interest that offer several
	mechanisms to support spatial random access or region of interest
	access at varying degrees of granularity. It is possible to store
	different parts of the same data using different quality.  This allows
	us to place the lowest frequency pieces in the fastest storage, and the
	highest frequency either on the slowest storage tiers, or if the data
	sizes are prohibitively  costly, not even write out the highest
	frequency. This binning of data in frequency space can take as an input
	(storage bandwidth knowledge, storage size limitations, and user
	intentions of how long the majority of the data will be frequently
	accessed.  In order to fully take advantage of frequency based
	re-organization we have devised a scheme where the data pre-conditioner
	first sorts the data in bins of 1024 words, and then we maintain the
	sort order, and then we used wavelets and spline fits to then
	re-organize the values.  This re-organization allows us to understand
	that the resultant data is smoother now, and can be reduced in each one
	of the bins.  \end{itemize}

@@

We will also research the time it takes to re-organize the data and use the
multi-layer approach to staging to understand where the data gets reduced and
where it gets re-organized. Open questions are: 1) do we move the data over to
other nodes to then reduce and re-organize, 2) do we first reduce the data and
then later re-organize this data on different nodes?  3) do we use different
cores on a node to do data-refactoring.  Similarly we will examine the cost of
reading the data. It is clear that if we only read in the high-importance data
it should improve the overall time to move the data from the SSIO layer to the
application, but what happens when users want all of the data? There is an
obvious penalty of having to take data in a hierarchical layout and then place
this back to a single-layer view.  Additional research questions which must be
addressed: Will some types of data (PIC data, finite element data, ...) be
conducive  for data-refactoring than others? How much semantic-domain knowledge
will we need to get a dramatic improvement of the storage system over not doing
anything special?

@@

\paragraph{High Impact Challenges:} Maximal data reduction on the SSIO layer
would result from storing only the code and the input parameters, and simply
recalculating all of the results. This extreme is not feasible in practice
owing to the size of the computation involved, but sheds light on the path to
achieving a step change in data reduction. We shall investigate the possibility
of writing a minimal set of information, and placing an ``auditor'' code in a
docker container so that data can be re-generated with a {\em local}
approximation of the full simulation. This is an entirely new approach to
tackle the data overload by using preprocessors in which information is not
simply stored in the form of raw data, but rather in the form of appropriate
parameterized algorithms. A priori information on the nature of the data is
used to choose the basic algorithm while the actual values of the data are used
to tune the parameters used in the reconstruction procedures. We shall refer to
this paradigm as an Auditor.  We will examine the use of several auditors for
the simulations which we will work with. One approach is to use a fluid-code to
audit a kinetic code, which can be a good approximation over short periods of
time. Once there is a large deviation (specified by the user) from the auditor
to the original calculation, the data will be saved. This is a local operation,
since the deviations will grow differently in space and in time.
